\documentclass[a4paper]{article}
\author{Flammable Penguins}
\title{\textbf{Jewel of Worlds} \\ Concept Document}
\pagestyle{headings}
\begin{document}

\newcommand{\TinySection}[1]{\begin{flushright} \textbf{#1} \end{flushright}}

\maketitle

\tableofcontents

\newpage \section{Introduction}
Picture hand crafting a select squad of units, a fine balance between tick and tack. The best and most suitable warriors from a range of eight playable races all with different mechanics, history and feel. You step through a portal onto a lush terrain which you've never seen before and begin moving pieces like a great general playing an overlarge game of chess.\\
Spells fly, swords clash, and arrows rain down in battle. At a crucial moment oyu use the correct spell or a clever back tactic the enemy didn't see comming. Then finally you wipe out your foe in a blase of glory. You step through the portal again, some for victory and others to return home.\\
Welcome to the Jewel of Worlds a fantastical turn based stratergy game, the likes of which you have never seen before.

\newpage \section{Background}

\subsection{The creation of the world sphere}

Eons ago before the gods were made and the energies and powers of the worlds were raw elements, the stuff of creation, the twelve worlds were created and put apart from each other so that peace may be kept. The power could not be in all twelve worlds so the world gem was created, a jewel beyond comprehension. It would contain the awesome power of the universe too frighteningly beautiful and full of power to behold. \\
The jewel of worlds was a beautiful jewel which shifted from the ivy emerald to scorching firestones as those only found in the Fires of Jur Rin Teerol. The only constant factor of the jewel, other than its potent contents, was its faceted portals. Each woven into an integral web mantaining each of the twelve worlds in their seperate realities, as to mantain the balance. Each world was unique and their gods different, some such as the Holy Worlds of Ireekal, where the almighty resided. Then there was the alchemical goblins of the warens of Gu'rak had their idols of gold and silver.\\
All was fine for eons as the very existance of other worlds, and knowledge of the jewel, were lost to most. Peace rained in the twelve spheres and all was good.

\subsection{The Fury of the Spectars}

On one world the memory of the jewel was fresh and never far out of sight. The spectral world of the Phantoms was a weak point in the jewel. The power was evenly distributed to all creatures of the world of Phantoms, and as a result they were all a step away from the portal, with no guardians at the gate.\\
As a result the flawed reality was dull and the specters soon started to visit the other worlds. They haunted, possesesed and interfered with the other worlds. The peace of the worlds was fractured ever so slightly, some more affected than others. The monks and holy warriors took them as an opportunity to learn more about the greater deity. The Daemon world of the devils and hellfire creatures, however, was quick to declare war upon the intrepid visitors.

\subsection{The Daemon Mage}

A powerfull lord of the brimestone realm was a mage by the name of Azeroth. He studied ancients arts, lost to most of his race until he found some fragments of the Song of Creation. Using a few notes of power and some defiled knowledge of the ancients, he created a spell which he though would seal his world from the invaders. In truth he was a hero of his people attempting to save them from a war, but in his folly he attempted to improve on the work of the gods.\\
Through many nights and many years, did he work endlessly to save his people. In a tower wrought of skulls and blood (the only material he knew to conduct magical energy) he focused the energies of the world to his great work. Endless selfless toil paid off in trumps as he attempted the ultimate folly of correcting the work of the gods.\\
Some say he would have succeeded; others doubt his magic would have done any good at all, but the truth shall never be known for the politics of the realm of fire are complex. The majority of support however, went to finding a way to wage war across the barrier, rather than sealing it. The war party laid seige to his keep and sought to undo his work. On the final night of cantation, as the last lines and sylables were being read out, the keep fell around the mage's ears. He hurried to complete his work before the final moments. Alas, the door burst open in a firey assault and a dagger wrought of silver and stone pieced his heart. However great magics cannot be undone by force, and the spell twisted and turned as the War mages attempted to contain it. But the workings were too far gone and the spell broke past and incinerated almost all the life from both the daemon world and the world of the spirits as it ripped across the jewel of worlds. The power hit the one weakness in the jewel, that of the spirirt realm. The jewel shattered.

\subsection{The Breaking}

The jewel slowly cracked. Some say it took the period of a eon and others say it happened in a heartbeat; the truth is, the time it took for the jewel's essence, the creator, to shed a tear was the time it took for his work to be undone.\\
The fragments sprayed across the dimensions like daggers of choas and disorder. Fragments rained down on the Holy cities on the mountain peaks like a furious wrath. The Warriors' noble combat was torn as javelins of disorder rippled the battle field as earth turned to syrup. The desert Mages saw the skies bring the rains early, as the very world drifted down through the bottleneck of reality, breaking like a million grains of sand. The Mechana watched as their great cities and endless works of creation were torn down by earthquakes of unkown magnitude.\\
The never-dead world of mog, mist, marsh and swamp bubbled and boiled as ice engulfed them; the shattered the endless bogs. The monstrous and bestial ember world of volcanic ash and violent quakes stood still as the skies cleared and the shamans paused for the barrage of unforgiving malace. The alchemists' labs exploded and their caves fell in as the sky became visible through cracks in the ever-roof of the caverns of Dirokeeel. The forests sighed as somewhere a child cried for the loss of beauty.\\
The few spirits and daemons not incinerated by the first bolt of terror fled through the now visible holes in their lost worlds. Forever without companionship or a homeland. Terran shook as order collapsed and the dark ages fell upon them. Thankfully at the time of the blast the creator had his attention turned upon Terran soil and shielded the Earth from most of the blast. The world of endless beauty which all have heard of but none seen. The world first created and said to be the home of the perfect race was cast out into the dark twisted purple ringed spirals of the broken realities never to be seen again.\\
The eight worlds which experienced the breaking were doomed but in a last act of love the creator shielded his children, creating an ether which would hold the fragments in order and allow life to survive. The space of eternity was replaced by the ether and the worlds which exploded into one great expanse now float as little islands through the endless ether. The creator slept.

\subsection{The Portal Wars}

Soon after people recovered from the breaking, counting their losses. They soon saw, through the misty domes which contained their floating homes, other floating islands. Some looked familiar but most were strange. Astronomers were created for the first time and slowly they realised they were not alone.\\
Shortly afterwards the inventive alchemists discovered the holes appearing on their islands were portals, after randomly jumping into some like the reckless folk they are. The goblins explained to the other races that the portals were like clouds; they appeared and dissapeared. They studied them and soon portal travel was a science, for the most part. People could begin to travel between the many islands and interact.\\
This first contact soon brought forth tensions and it was seen that the races would need to meet. So they found a fragment of the jewel which was neutral territory and established the council chambers, so that all races might meet and discuss their worlds. They anchored the portals to this place and the very nature of the jewel fragment prevented any form of violence or ill act, and it became the perfect meeting place.\\
Quickly the worlds involved in the breaking met for the first time and agreed upon three things. The first was that no act should ever be attempted to repair, alter or in any other way affect the order of the multiverse. The second was that no tool that destroys land or harms space may be used. The third was that Earth should be left alone, as well as the lost world. They do not know this sorrow and it would be cruel to bring it upon them.\\
Soon after these rules were agreed upon however, every race realised that the worlds were lost and that the three things people needed to live was land, food, and power. The only way to attain these things would be the reclaimation of the islands. The portal wars broke out, and once again war raged for the oldest reason, land.\\
Currently, the council still holds and the worlds still discuss, but open war rages on and the costs are high. The worlds are too different to co-exist peacefully. Peace was the very reason why the creator placed them apart, a fate which has been undone.\\
So now across the ether small armies meet in battle to decide the fate of the islands.

\subsection{Dance to Other Places}

You go to this new-age party at this freaking Scotish Castle; its an underground bash which only the cool cats know about. Despite being grounded you sneak out with some friends to have the time of your lives. You get to the party which is real great until the ground shakes and the power goes out. The rest of the party happens by candle light while the bands play unplugged. You fall asleep on a cold rock.\\
When you wake up you are in a sack being carried against your will. You panic as you hear grunts and rough sounds you don't understand. You manage to poke a hole in the bag and you see a volcano spouting ash into a red scorched sky which seems strange. Even wierder you see monstrous figures of the like from story books. You soon realise this isn't a dream, or nightmare.\\
Just as you start to panic you hear a clear silver trumpet ring and the thunder of hooves and clank of armour. Battle rings about your ears as you are harshly dropped to the floor and forgotten. Soon after what sounds like a victory for the new force you see a knife cutting open the bag. A handsome knight takes you up into his strong arms. His clear voice says words of comfort which you don't understand, as he steps through what can only be described as a shimmering hole of light.

\subsection{Game Premise}

Sylvia (the main character), along with several other humans, are pulled into the Portal Wars by unkown forces. At first the game is a quest to get home, but you soon get entangled in the politics and streams of the world. Soon your quest is to find the truth.\\
The story relies heavily on preconcieved notions and rethinking first impressions. You realise that the bad guys are not the all evil and the good guys are not all pure. But above all, you realise this crazy place is just filled with different people who want to go on living.\\

The story will be somewhat linear but it plays through the various races as you come to understand the different sides of things and sometimes you are just pulled in directions, against your will. You notice how your fellow earthlings react to the same scenario, and in a sense, dispite the linear nature of the story, we hope to convey the multiple possibilities through the different people who were pulled through with you. Their choices being different from the player's, the player will hopefully feel as if the different avenues are explored, even if not by him personally.\\


\newpage \section{Technicalities}

\subsection{Time management policies}

We understand this project has the unusual limit of having less than a year of developing time, with a team of three people. We also realise we may have a concept which appears at times as over-stretched. We feel we will meet all of our goals, and to explain our time management policies and resource management, we have added this section.\\

Simple reusable code structure is the key component of our game. We expect to have a workable engine which is playable if not pretty or complete by the end of March. From there it is a simply matter of populating the game and adding a few small code expansions.\\
The basic battle engine will be outlined in this document, and our concept document fully details all the units and their abilities, allowing us to measure how far our engine must go and the complexity which needs be incorperated.\\

Modeling shall be accomplish by using simple low-poly cartoon-like models. This adds to the storybook feel of the game and significantly cuts down modeling time. The low detail however, will not mean the graphics need to look poor.\\
We intend to code a set of effects and particle engines which will allow us to quickly, using simple maths, spice up and breath life into the simple models. Time-wise this is much more effiecent as we have upwards of 60 units to model, not to metion terrain objects.\\

FMVs and cutscences for the most part will be scripted events in the game world, which, with this engine, will be quick and easy to write. Larger cutscenes will be done in a pictorama style using a set of stills and simple transition effects, combined with an effective voice over, keeping the workload and file sizes of the game relatively small.\\

Balance issues will become apparent very early on. For this reason we are currently aiming for a multiplayer-only beta to be released at ICON; the biggest South African game-playing convention. From there we hope to attract enough attention to have our game played until it breaks, on a daily basis, and we will also have loads of feedback. This feedback will be vital to the successfull polishing of the game.\\

In a worst case scenario of deadlines looming we will make the desicion as early as the end of the semster as to what to cut from our game. The things listed in priority of cutting are as follows, starting with the first to be cut items.

\begin{itemize}
\item Conquer Mode
\item Story Dialogue Recordings
\item Extended Balancing
\item Shortend Story
\item Less Hero Units
\item Online Beta Testing
\item Terrain Sets
\item Internet Play (in other words only hotseat and network)
\item Unit Sounds
\end{itemize}

We truely believe, however barring any major upsets in our devolepment team or the process of this year, that we will achieve all things laid out in this document as planned. We are all very ambitious and believe that the project is the defining moment of our degrees as all of us want to become game developers.

\subsection{Technical Specifications}

\subsubsection{Open GL}

We plan to use the default OpenGL libaries supplied by card manufactures with the possibility of using some extended libraries should the need  arise. We plan to keep the game portable to Linux or other platforms. We will be using Windows and nVidia specific tools as they are our main focus, but we plan to implement wide band support on this front.

\subsubsection{Particle Engine}

We intend to code a fully functional particle engine with pixel shader support. This will be vital to making our world look attractive without loads of content generation as particle scripts will be made simple and quick to implement.

\subsubsection{Input}

We plan to use a fully supported mouse interface which will not require any keyboard input. The interface will also offer every option via the keyboard so the player can blend interaction between the two devices without difficulty. The interface will be focuses to a complex context with simplified focus. This means that there will never be greyed out or unavailable actions on screen and the screen will be uncluttered. We will be using the default Win32 input methods, hopefully with Linux alternatives. The removed input architechture we hope to use will also allow easy implementation of tablets and other third-party control systems.

\subsubsection{Multiplayer}

We plan to use direct multiplayer options using sockets and other low level technologies to keep the bandwidth requirements very low and avoid DirectPlay or similar dependencies which would get in the way of porting options. The backbone of the system will use a removed input system so as to implement the multiplayer or distance input seamlessly. The multiplayer modes available will be server and peer to peer. We also will implement a hotseat option.

\subsubsection{Map Editor}

We plan to build a fully functional Win32 map editor. Porting options will be limited but we plan to open source the file format. The map editor will be fully capable of creating maps or scenario (story) events. The display engine will be a default Win32 interface.

\subsubsection{Random Map Creation}

We are also planning a vital piece of the game for replaybility to be the a random map generator. It will be able to create a fully functional 3d map without human interaction. This will require a primitive construction AI which will be able to build a playable map rather than a messy collection of random values. The random map generation is key to replayability.


\newpage \section{Game Modes}

\subsection{Story Mode}

The game will use a linear story as we feel it promotes a greater sense of an epic tale, and gives us a more realistic goal to reach. The characters can be defined in preset monologues. The story engine will progress regardless of the player's wishes. The individual battles will however have an impact on small dialougue lines and future battles, as will be explained later in this document.\\

The story starts with a short FMV (pictorama), as described in the story section\footnote {Refer to background section for more information.}. Further story advancements happen between scripted cutscences (in game). There are also more FMVs (pictorama) and cutscences (ingame). The story throws you in with a small team of warrior units.\\

Before each story battle, you have a squad selection screen where you choose the units available in your barracks up until the army point limit. This is explained in game by the limit of the island portals which can only take so many people and artifacts.\\

You then fight a battle much like a skirmish or multiplayer game, aiming for your goals. At the end of every battle you see a debriefing screen were cutscenes happen, and the story is furthered. One or two lines of dialogue might alter according to your actions, but the general thrust of the story line continues.\\

What does get affected however is your barracks. If you loose a unit he is lost forever, unless he is hero in which case he would have been healed up. Other things likes large losses or cross-team combinations, anything which causes poor morale, can cause the shrinkage of the barracks. Story events and good morale will restock the barracks, or sometimes strip them empty and give you a fresh start, like when the story forces you to change teams.\\

We hope to convey the option of choice by showing several other humans in the story who knows the main character. These people make different choices and as a result the player is exposed to different story arcs.\\

\subsection{Conquer Mode}

A game mode which is last on our wish list. It is the pinacle of multiplayer or single, non-story, combat. Inspired by Heroes of Might and Magic and similar titles, conquer mode is a different out-game engine.\\

The game starts with you owning a base island with a few unit-producing facilities, the only resource being power. To gain power for building units and upgrading things you need to claim islands. Some islands also give you bonuses and different units not from your city.\\

You create armies of a certain size which wander a map of islands linked with different portals, some of which have time limits, army size limits, direction limits or phases. Slowly you want to claim more land and cities, but there are eight other players in this space.\\

Basically you fight until you control the entire map. You take over new cities and upgrade your current cities using power. The city structure is simple and generates units on a time basis which can then be recruited with power points. Your armies in some cases improve, your heroes might gain artifacts or new abilities.\\

Ultimatly it adds a whole new dimension to the game and for these reasons it is last on our list of things to do. It is realistic, but we are banking of having a major upset which stops us finishing this feature, in which case we will code it in after the marking peroid.


\newpage \section{Battle Map}

This document explains the process behind the Battle Map system. We will try to document ideas already explored but they are numerous and disjointed so for the most part we will focus on what we will be using.

\subsection{Needs}
\begin{itemize}
\item A 3d map with objects support, in other words houses and trees along with cliffs and caves.
\item Small filesize, as to aid multiplayer support.
\item Constructable block as to enable random map generation.
\item Tile swapping, for ease of Variety.
\item Low Content requirements and interpretation of data. Maps must be quick, things like slopes and stuff must be calculated by engine.
\item Realtime Terrain, as to allow for the destruction of objects and recalculation of lightmaps ect...
\item Grid Base
\item Fast Checks, the map format must not slow down the engine with complex calculations
\end{itemize}

\subsection{Theory}

Taking into account our needs we came up with the following. The main part of the map is the ground. The ground can be represented as a grid with each grid piece being at a certain height and of certain properties. Taking this into account you can accurately represent the ground with a height map. We can also say that the height of a single map will never be more than 64 units. Therefore each point of the map can be stored as a 64bit piece of information, in gametime it would be a short as its quicker.\\
The problem with a height map is that its not true 3d. We solve this with 'objects'. We have a collection of what can be best described as objects all over the terrain. Things such as houses, trees, rocks, overhangs ect... Each object is a display of a textured model. In game terms however it is represented by a voxel tree. The best way to describe this is a bunch of lego blocks. We can safely say that over half of the blocks will be 'open', in other words, have no objects on top of them. We can also safely say that most objects will not even take up a quarter of the space between the sky and ground. So therefore we need to use a storage method that is fast and not space clogging. So we use a stack.\\
Each object voxel tree is broken up into columns and then stored onto the base of the map. Now while it would be more efficient to use a block system of each, grid voxel being a solid block, it would work because of walls and other thin objects. Therefore each voxel piece is actually three walls of a cube. The reason being is that the cube is surrounded by cubes and to have six walls would cause redundency.

\subsection{Building of Objects}

All objects have two pieces, the first one is the graphical representation of the object, for instance a tree. This is a relatively high-poly and pretty object. The model is skinned and any other graphical mods are applied.\\
The next and most important part is a logical representation of the object. For example the tree is broken into cube walls, or in some cases solid cubes. So we construct a logical object, as solid and hollow cubes.\\
Then the map editor or randomisor builder takes the logical object and drops it onto the terrain and dissasembles the object into columns of data as discussed above.

\subsection{Quick Reference on Flags}

The following are just for clarity. A flag system is very fast because the binary AND and binary OR operations are its only tools. It is small and efficient. Lets explain, all numbers on a PC are base two and boolean flags are base two numbers in essence. Let us take a cube as used above. Example.
Cube Needs to store three walls and health. 

\newpage \section{Description of Play}

We shall now describe how skirmish or multiplayer system works. The story and conquer game modes overlay the skimish mode in such a way the exact same battle engine is used.

\subsection{Choosing of Map}

The selection of map is made by the host and then agreed upon by the other player. You will have a choice between a premade map or a randomly generated map. The map tileset must also be set at this point. The current map tilesets are as follows.

\begin{itemize}
\item Forest		\' Nature
\item Caves			\' Alchemy
\item Volcanic		\' Monsters
\item Swamps		\' Undead
\item Foothills		\' Mechanical
\item Desert		\' Mages
\item Plains		\' Warriors
\item Mountains		\' Holy
\end{itemize}

At the current time the teams gain no advantage from terrain but that may change. Once map selection is done we move onto scenario selection.

\subsection{Scenario}

The default scenario is skimish but other option are available as described below.

\subsubsection{Skimish}

Each team has an equal amount of AP and a small deployment zone at opposite sides of the map. They deploy simultanously and play commences. The winner is the last man standing.

\subsubsection{Kill the Leader}

Same as Skimish, but the last hero standing determines the winner.

\subsubsection{Ambush}

One player gets a large army and the other player a small army. The small army gets to deploy in any location on the map and all units start fully charged and deployed. The large army then deploys in a small area as per skimish. The play continues until the last man standing is done.

Other play modes considered are capture the castle, siege, delivery, and rush.

\subsection{Army Building}

You must then spend your army points on units. Each unit has a set cost. Any unit from an allied team costs ten percent more. Any neutral team costs fifty percent more. No foe or enemy team units may be selected. Only one hero of each name may be in play. More varied teams have weaker morale.

\subsection{Deployment}

Depending on game mode your deployment may change. Standard deployment is as follows:\\
 The player can deploy on a small area, roughly a tenth of the map, on opposite corners. Units are deployed in any order. \\
 Initiative is decided. This is achieved by adding all the units together and evenly deploying the units by team randomly. What this means is if both teams have five units team A moves a unit followed by team B moving a unit. In the case of imbalance in team numbers the units are evenly spread as to minimise the weight between turns. Example: 00X00X00X or X00X00X0X00X0. The order of units is random but the controller is not.

\subsection{Commencement of Play}

From this point on we choose the next unit on the intiative stack.

\subsubsection{The Maintance Phase}

During this phase all start-of-turn events occur and recharging or powering up.

\subsubsection{The Action Phase}

During this phase the unit may perform the following actions until he chooses to end the turn or runs out of time units:
\begin{itemize}
\item Walk
\item Jump
\item Climb
\item Swim
\item Turn
\item Crouch or Standup
\item Attack*
\item Deploy*
\item Cast Spell or Use Ability*
\end{itemize}
\footnotetext{* Not all units have these abilities}

All actions require time units and some actions require other resources.

\subsubsection{The Clean Phase}
Any spell effect or limited effects due to end this turn are ended.

\subsection{Conclusion of Play}
Play continues until the win conditions or lose conditions are flagged.


\newpage \section{Magic of the Teams}

One of the diverse factors we wanted to implement was magical casting and how it worked differently for each team. The fantastical setting allows us this freedom. It is important to note that hero units may use completely different system from that of their teams. For instance, Merlin uses the mage system but is part of the warriors team. Also note that magic and abilities fall under different categories. Some units have the conceal ability or other non-standard actions which just use time units.

\subsection{Warriors \\ Long Turn Casting}

The warriors of old are fashioned pretty closely to medievil European combat. As such they have no 'magic' units as such. What they do have are sappers or engineers. Each engineer gets two toolsets, each of which supply him with the tools for an engineering task. Some of the sample tasks are listed below. There is no limit to the amount of uses a kit has; the catch is some of his more complex tasks take several turns to complete.

\begin{description}
\item[Barricade], building a bardicade which allows pike men and archers to have an extra layer of defense.
\item[Pit Trap], builds a trap which will fall an infantary unit, taking all his remaining TUs and dealing small amount of damage.
\item[Track Wreck], makes a small area of terrain difficult terrain
\item[Ladder Drop], gives him the ability to drop a rope ladder.
\item[Bridge Builder], can bridge large gaps.
\item[Undermine], collapses a cliff, causing a damaging avalanche and creating a ramp.
\item[Clear the way], allows destruction of obstacles and clearing of rough terrain.
\end{description}

\subsection{Mages \\ Mana List}

The mage team has several casting units but they all use this system. The mage fills a spell book with spells. Each spell has a power rating. The mage is limited to selecting either a broad range or a specialised range. For example a mage could take 8 power rating 1 spells or one power rating eight spell, or a mixture.\\
The mage starts with a set amount of mana, but there are units that can help recharge mana for a mage. The mage can cast any spell numerous times, but each time he expends mana. The mage may cast as many spells as his TUs (Time Units) and mana allow.\\
The spell list is too diverse to list here (see team unit descriptions for a full list), but their magic falls into three catergories:

\TinySection{Vulgar}
These are direct Effect spells which attack a certain target with a certain effect. E.g. Fireball

\TinySection{Summon}
These spells bring forth entities to fight for the mages. E.g. Earth Elemental

\TinySection{Strategic}
These are indirect spells which give the mage a long term stratgic adcantage. E.g. Mage Armor

\subsection{Mechanical \\ Expensive Abilities}

The mechanists have steam pressure statistic which will be full at the start of every turn. Each unit may have one, two or possibly three powerful abilities, which use TUs and Steam. But for all intensive purposes the steam is just there to ensure not all the TUs are spent on abilities.\\
The abilities will be very mundane offensive, defensive or ranged attacks. Ranged attack will also use steam. For more detail see the unit description in the Teams section.

\subsection{Undead \\ Death Magic}

For every unit that falls more shall rise. The undead magic users draw strenght from the power of death. The amount of power is determined by the unit's AP (Actions Points) and proximity to the mage. However, because of the frailty of undead mages, being in the thick of battle is dangerous for them. Also, any unit killed by death magic does not release mana.

Three Spell areas:
\TinySection{Reanimate}
Create undead units on the field of battle.
\TinySection{Cripple}
Damage enemies with harmful effects.
\TinySection{Boost Undead}
Aid undead units.

\subsection{Monstrous \\Deck of Cards}

The shamans of the beast and monster races are powerful but unpredictable. Each magic caster has a figurative deck of magic cards, of a minium size. The spell has to be used the turn its drawn or it is burnt. So the magic of the monsters is a gambler's tool and very powerful or very weak depending on your luck.\\
The spells are drawn from all other races.\\
Also mages can possibly have a hand of cards as opposed to a single card.

\subsection{Alchemists \\ Expendables}

The brewers of potions, magic dusts, and exploding toys. Alchemy is an art which cannot be praticed on the field of battle. However, the potions and dusts concocted can be carried into battle.\\
Almost every goblin on the field of battle has a potion belt which holds the potions of his people. He hurls them much like grenades. Whenever a goblin dies there is a chance of his potions going kabooom.

\subsection{Nature \\ Recharge}

The nature of the forest is that of patience. Every nature mage has magic which, when cast, will slowly refresh, on a time basis. The system allows him to use his spells in any order as long as they are refreshed.\\
Nature mages will also have a meditate option to refresh faster while doing nothing.

\subsection{Holy \\ Passive}

The all-embracing one is everywhere and nowhere, his voice is a whisper on the wind. As such prey will bring passive auras of strength and slow effects. There are no magic casters in the holy city but the priests and holy men can draw strength from prey and inspire those around them.


\newpage \section{Unit Creation}

In order to streamline the creative process of units we have devised the following stratergy for the unit creation pipepline.

\subsection{Concept}

The first and possibly the most challanging phase was when we created the idea of the unit. This includes its past, attitude, game stats, look, and general playability.

\subsection{Artwork}

We then create concept art and retrieve reference material such as photos or other real life references. All these are used to further flesh out the character and give the modeler something to work on.

\subsection{Modeling}

The modeling process is the first place reusability comes in. Many units have similiar looks or components so the first step of the modeler is to find any of the previous models which he can use to save time, as its an important resource. The modeler completes the model and skins it. Prelim Skins are also created at this point.

\subsection{Animation}

The character rigging will take place here with biped whenever possible as to be able to reuse motion scripts to streamline the animation process. The required animations for all characters are as follows.

\begin{itemize}
\item Walking
\item Idle Animation
\item Attack Animation
\item Special Animations (casting, throwing, deploying)
\item Crouch idle animation
\item Crouch Attack animation
\item Crouch special animation
\end{itemize}

At this point the models are exported into md2 format. We shall use third party tools for this.

\subsection{Logical Creation}

The unit is then created by expanding the most revelant parent unit, and adding the new features of the unit. The philosophy of code reusability is core to the coding structure in order to streamline the unit creation process.

\subsection{Testing}

Extensive game testing will be performed by outside parties and fine tuning will constantly occur to units. These changes will be quick and easy to implement thanks to the coding structures in place.


\newpage \section{Unit Anatomy}

What makes a unit tick and fleshed out.

\subsection{Concept}

The concept and general character holding the unit together is the most important part of every unit for us. Each unit is firstly a character and secondly a game piece.

\subsection{Graphics and Sound}

The look and feel of the character is defined by the model, skins, animations and very much the sounds of the character.

\subsection{Base Unit}

The gameplay of the unit is vital as no amount of character will make the game playable, the character only makes it enjoyable. This part defines the vital stats as described below:

\subsubsection{Primary Attack}

\TinySection{Melee}
\begin{description}
\item[Accuracy]: Determines the chance of hitting someone.
\item[Damage]: How much damage is dealt
\item[Reach]: Is the weapon a reach weapon or not
\end{description}

\TinySection{Ranged}
\begin{description}
\item[Accuracy]: The amount of deviation over distance.
\item[Damage]: How much damage is dealt
\item[Reach]: The amount of distance before the projectile drops. Note that  this is not the maximum or minimum distance. Shooting onto higher terrain or lower terrain changes your range. Your reach is constant however.
\end{description}

\TinySection{Indirect}
\begin{description}
\item[Accuracy]: The amount of deviation
\item[Damage]: The amount of damage dealt at ground zero
\item[Splash]: The size of the blast radius.
\item[Reach]: The distance the shot can go, affected by terrain height or source, not target.
\end{description}

\subsubsection{Secondary Attack}

See Primary Attack, most units do not have a secondary attack.

\subsubsection{Defense}

\TinySection{Dodge}
\begin{description}
\item[Melee]: Your chance of avoiding taking damage
\item[Range]: Not aplicable
\item[Indirect]: Your chance of taking half damage
\end{description}

\TinySection{Armour}
How much damage gets soaked.

\TinySection{Health}
Your total amount of hit points before you die.

\subsubsection{Movement}

\TinySection{Speed}
How much moving across terrain costs you in time units.

\TinySection{Agility}
How good is your jumping, climbing, swimming and other athletic feats.

\subsubsection{Morale}
How brave are you before you flee combat

\subsubsection{Time Units}
Your total amount of time units. For most units this will be constant.

\subsubsection{Army Points}
How much this unit costs

\subsection{Common Extensions}

Some common extensions are listed by team:

\subsubsection{Warrior}
The sapper has a list of abilities which all have extended time unit costs.

\subsubsection{Mage}
All mage units have a power rating, a spell list and mana. Power rating is how much mana they get each turn and mana is how much they have stored. Each spell has an effect and cost in time units and mana.

\subsubsection{Mechana}
All mech units with the exception of the Steamhead have a steam stat.
All mech abilities have a steam cost as well as a time unit cost.

\subsubsection{Undead}
Undead mages have a life force stat which records the amount of death they have stored. They also have spell lists.

\subsubsection{Monsterous}
Card casting is a system which needs to store a deck of say fourty cards, randomly sorted and all referenced to. 

\subsubsection{Alchemists}
Most goblins have a climb bonus.
All goblins cannot swim.
All goblins have expendable potions which have an effect, referenced to, and an army point cost.

\subsubsection{Nature}
All nature casters have spells which each have a refresh timer. Each spell has its own timer. The effects are not stored in the unit, but only referenced to.

\subsubsection{Holy}
Strike Undead is a double damage modifier which some units use.
Auras have range and effects, referenced to, which are applied to units in the area of effect depending if they are friendly or hostile. Auras are always active, one at a time per unit capable.

\newpage \section{Tertiary game mechanics}

\subsection{Morale grid}

Concept is quite simple, picture a octagonal shape. At each vertex is a team. The closer the team the better the teams get along. For example, take the Monsters. They work well with the alchemists and the undead. They are neutral towards the Nature and Mechanical teams. The fight the Holy and Mage teams and outright despise the Warriors. This means that if you have a monster team you can have both Alchemists and Undead in your team with little or no problems.\\
You can also have Nature and Mechanical units in your team but you can't have both mechanical and nature because they are mortal enemies. And if you have Nature and Undead then you are going to get penalties. In other words the closer you come to a half circle the weaker your team moral and bonus are.\\
You have too make a choice between a highly focused army and specialised army with good morale and team bonuses or a wide range off units with lots of skills and abilities but a weak team morale and spirt, or somewhere in between the two extremes.

\subsection{Weather}

Bad weather is bad for everyone. Some units are less affected than others. For example flying units are the most affected by adverse conditions, the most extremes completely stopping them while the the lesser extreme severly hampers them.\\
Also things like climbing are badly modified during rain, while jumping and athletics are mogged down by snow.\\

Clear\\
Fog\\

Wet\\
Drizzle\\
Rain*\\
Monsoon**\\

Snow Covered\\
Sleet\\
Snow*\\
Blizzard**\\

\footnotetext{* Only Twilight or Dark}
\footnotetext{** No Day Cycle}


\newpage \section{Team Descriptions}

\subsection{The Warriors}

\subsubsection{Units Summary}

\begin{description}
\item[Sword Slinger]- Serf with Sword
\item[Pikeman]- Serf with Pike
\item[Captain]- Enlisted Swordsmen

\item[Bowmen]- Serf with Bow
\item[Tower Guard]- Efficient Crossbow wielder

\item[Ballista]- Single unit siege weapon
\item[Catapult]- Area of Effect (indirect fire)

\item[Sapper]- Support Unit with tools
\end{description}

\subsubsection{History}

From the ever flowing plains of the golden homeworld constantly at war, the Warriors step forth. The history of the world is littered with nobility squabbling over minor matters of state and throwing men into honourable combat for a quick resolution.\\
The king ruled over all, though he died in the Breaking. Since then the House of Lords has been the main ruling body and they have united against the common enemies. However, the squabling and belittling still happens between those of different provinces, even though it is mostly from old habit than anything else.

\subsubsection{Inspiration}

Midievil Europe, during the War of the Roses.

\subsubsection{Detailed Unit List}

\TinySection{Sword Slinger}
Low end melee unit without any special abilities. Decent attack and low defense due to lack of armour and shield.
The unit's vocal response and attitude is that of a repressed serf. He is extremly cynical.

\TinySection{Pikeman}
Low end unit with a reach weapon, most effective when they are behind defensive or friendly units. They cannot attack targets next to them. Again the unit is a repressed serf with cynical approach, and he is convinced he is going to die first, as the token nameless peasent always dies first.

\TinySection{Captain}
High end melee unit, good defense and attack. An honourable unit who constantly says things like, 'For Honour and Death' or 'Into the Breach my friends'.

\TinySection{Bowmen}
A long range, high damage indirect fire unit. Accuracy is not great and the defense extremly poor. Melee units cut down the bowmen. The bowmen are serfs who have escaped the melee. Attitude is very much cowardly.

\TinySection{Tower Gaurd}
A extremly good short ranged unit with a backup melee weapon. The unit's defense is good. They follow the 'Defend the Keep' philosophy and say things like, 'Bar the gates' or 'Man the parapets'.

\TinySection{Ballista}
A good siege weapon. Its direct attack is extremely powerful, its movement and defense are low and it takes a turn to reload. The ballista is pushed by a single serf, who complains constantly about the weight.

\TinySection{Catapult}
An indirect weapon which rains death upon people from afar. Its accuary could be a bit better and its slow. Its range, however, is extremly great. The catapult is pushed by two serfs who are constantly talking calculus.

\TinySection{Sapper}
A support unit which has a toolbox. No attack stat but defense and speed is good. The unit relies on being able to run away or be saved by other units. The toolbox can be filled with several options. The attitude is professional and overconfident.

\subsubsection{Battle Engineering}

The warriors of old are fashioned pretty closely to medievil european combat. As such they have no 'magic' units. What they do have are sappers or engineers. Each engineer gets two toolsets, each of which supply him with the tools for an engineering task. Some of the sample tasks are listed below. There is no limit to the engineer's use of his tools, the catch is some of his more complex tasks take several turns to complete.

\begin{description}
\item[Barricade], building a barricade which allows pike men and archers to have an extra layer of defense.
\item[Pit Trap], builds a trap which will fall an infantry unit, taking all his remaining TUs and dealing small amount of damage.
\item[Track Wreck], makes a small area of terrain difficult terrain.
\item[Ladder Drop], gives him the ability to drop a rope ladder.
\item[Bridge Builder], can bridge large gaps.
\item[Undermine], collapses a cliff, causing a damaging avalanche and creating a ramp.
\item[Clear the way], allows destruction of obstacles and clearing of rough terrain.
\end{description}

\subsection{The Monsters}

\subsubsection{Units Summary}
\begin{description}
\item [Harpies]- A fast half-bird flying unit that does damage with its claws
\item [Banshee]- A hidious female shaman capable of devastating magic
\item [Medusa]- A female figure with snakes for hair and a petrifying gaze

\item [Cerebus]- A three headed hound with swift movement and a poisonous bite
\item [Minotaur]- Slow moving but viscious half-bull, half-human beast that can charge into battle
\item [Cyclops]- A one-eyed giant that hurls large chunks of earth at its enemy
\item [Hydralisk]- A seven-headed lizard that can attack on all sides
\end{description}

\subsubsection{History}

In the times before the Breaking the species that together were only known as the Monsters were at war with each other. One humanoid being, as tall as all the volcanoes of the land combined, known as the Titan, was the only element of control present. The Titan, although he rarely interfered with the ways of the beasts squabling underneath his feet, ensured that no single species became more powerful than any other. However, after the Breaking the Titan was no longer present. No being in the Realm of Monsters was aware of where the Titan has taken his stride. As chaos was about to erupt into utter domination of species over others, they also became aware that the broken lands were not big enough to host even one victorious dominant species, and cooperative expansion was the only available option to ensure survival of even only the fittest. The Warriors, seeking adventure, fame, riches and egoistic pride, are their greatest enemy. The Holy people wish to destroy all that will not bow to their masters, and the Mages derive pleasure from looking on as the Monsters succumb to their ordered magic. They shall not be untrue to chaos, and their enemies will never rule the beings that are Monsters!

\subsubsection{Inspiration}

Bulfinch's Mythology and many movies based on Greek and Roman mythology.

\subsubsection{Detailed Unit List}

\TinySection{Harpies}
Harpies have the upper body of a 14-year old with large claws at the end of each leg. The wings on their backs grant them the ability to fly at high speed. Their appearance is enough to make any babe weep and cause any adult restless nights. They nest high up on the volcanic hills of their homeland and their claws do damage to anyone foolish enough to oppose them. 'Let me tear flesh from bone!'

\TinySection{Banshee}
These female shamans vaguely resemble human females and move around the volcanic terrain at a high pace. They can do moderate damage to anyone they are hostile to. They also specialise in powerful magics instead of direct combat. 'Pick a card, any card!'

\TinySection{Medusa}
Although similar to appearance to the Banshee this unit has snakes instead of hair. This grants her the ability to spot enemies all around her. Should anyone find themselves unlucky enough to meet her gaze, they will instantly turn to stone. She is also capable of hand to hand combat and some minor magics. 'Look into my eyesss and tell you wissssh to be passsionate!'

\TinySection{Cerebus}
This three-headed hound crawls from the cracks in the volcanic mountainside and can move swiftly across the terrain. Its bite is poisonous to anyone failing to keep it on a leash.'Throw us your bones!'

\TinySection{Minotaur}
This beast has the head of a bull and the body of a human. Despite its slow movement it is well armoured. It can also charge into battle swinging it huge axe to devastating effect.'I strip flesh from bone and feed it to Cerebus!'

\TinySection{Cyclops}
As the cyclops only has one eye, it has difficulty to see what is not directly in front of him. Despite this he has no trouble traversing the volcanic terrain of its home. Because of his tough skin it is difficult to deal damage to. If annoyed the Cyclops will pull large chunks of earth from the very ground and hurl it at its enemies.'Let me see you play CATCH!!'

\TinySection{Hydralisk}
The hydralisk is a large four-legged lizard with a short tail and seven heads. It was spawned it the pools of boiling water amidst the lava of its volcanic homeworld. Because of its seven heads it is capable of seeing all that is around, and can attack anyone within range at any side of its body. It is fairly difficult to kill and does great amounts of damage. 'You cannot hide from Hydra!'

\subsubsection{Fear}

All monster units have a fearsome value. This is how terryfing the other races perceive them to be. The mere presence of monsters causes morale damage. The resulting effects can be panic, confusion, paralysis, and outright retreat.

\subsubsection{Shaman Magic}

The shamans and magic users of the Monsters use chaotic random magics, represented by a stack of tablets, simply called cards, or the card system. The shaman can choose to cast the spell represented on top of the stack, or can discard it, losing it forever. Because of the chaotic nature of shaman magic, they have access to all known spell in the known existance.

The available spells in the deck of cards follow:

\TinySection{Reanimate}
\begin{itemize}
\item Hopping Eyeball
\item Skeleton Trooper
\item Skeleton Archer
\item Zombie
\item Anathema
\end{itemize}

- Voodoo Curses
\begin{description}
\item[Ghoul Touch]: Damage target and steal life force
\item[Cripple]: Target looses half time units
\item[Despair]: Destroys morale
\item[Taint]: Makes an area of terrain or body of water damaging to the living

\item[Haste]: fifty percent increase in time units
\item[Rage]: Target unit deals double melee damage
\item[Regen]: Fully heals undead unit

\item[Regeneration]: Slow recuperation of health
\item[Courage]: Restores morale
\item[Honed Assault]: Improves attacks
\item[Shining Soul]: Improves defense
\item[Holy Message]: Delievers speed to the feet of warriors

\item[Ice Bolt ]: Low Damage, reduce TUs next turn
\item[Fire Bolt ]: Weak Damage (better than Low Damage)
\item[Lightning Bolt ]: Low Damage travels on

\item[Freeze Ball]: Explosion which freezes targets
\item[Blaze Ball]: High damage explosion
\item[Sonic Boom]: Damaging explosion which stuns

\item[Ice Shard]: Damages and freezes target
\item[Lava Blast]: Extreme direct damage
\item[Chain Lightning]: Moderate damage that jumps until it can no longer

\item[Haste]: Increase TUs for target, next 2 turns.
\item[Time Speed]: Increase Tus for team, next turn.
\item[Slow]: Decrease TUs for target, next 2 turns.
\item[Bullet Time]: Decrease TUs for team, next turn.

\item[Teleport]: Jump to any visible point on the map
\item[Recall]: Recall any ally to your side.

\item[Sonar]: Detect the location of all foes.
\item[Feather Flight]: Fly for a turn.

\item[Minor Elemental]
\item[Elemental]
\item[Major Elemental]
\item[Sand Golem]
\item[Iorn Golem]
\item[Rock Golem]
\item[Wall of Fire ]: Creates a damaging wall of fire
\item[Air Wall ]: Creates an impenetrable wall for one turn.
\item[Ring of Fire ]: Creates a damaging ring of fire
\item[Dome of Air ]: Creates a invulnerable dome for one turn
\item[Elemental Armor ]: Protects target
\item[Bridge of Air ]: Creates a bridge of air
\item[Minor Portal ]: Creates a portal for one turn to visible location.

\item[Mage Armor ]: Increases defense for the mage only
\item[Far Sight ]: Looks at a small area of the map
\item[Blink ]: Random short range teleport
\item[Mana Strike ]: Damaging melee attack
\item[Mana Arrow ]: Damaging ranged attack
\item[Leap ]: Allows a long or high jump.

\item[Concealment]: Unit obtains invisibility while stationary
\item[Blind]: Target loses half time units

\item[Summon]: Grizzly Bear, Hawk, Fox
\item[Tremor]: Area effect that stuns and damages target
\item[Bog]: Target loses half time units

\item[Tree Stride (Nature Stride)]: Teleport between two lands objects
\item[Entwine]: Target cannot move next turn
\item[Bark Armour]: Increases defense
\item[Poison]: Damages target over time

\item[Water stride]: Teleports between two points in a body of water
\item[Seduction]: Target loses half time units
\item[Heal]: Heals injured unit by fifty percent
\item[Mist]: Area effect that reduces sight range
\item[Encourage]: Morale increase
\end{description}

\subsection{Holy City}

\subsubsection{Units Summary}

\begin{description}
\item[Apprentice] - Basic melee unit
\item[Kai Monk] - Fast well trained melee fighter
\item[Healer] - Support unit with brilliant healing
\item[Reliquery] - Aura generating box carried by two apprentices
\item[Slingshot] - A powerful direct fire weapon

\item[Templar] - An extremly powerful reach melee warrior with low defense
\item[Samuri] - A well balanced and powerful melee unit
\end{description}

\subsubsection{History}

The holy cities were built eons ago in the high mountain tops so they could be closer to God and his domain as they believe him to be above all creation. They sought to be pure in life so they might ascend to the heights above.\\
The head of the order who was a wise and enlightened soul passed away shortly before the first council of worlds. His guidance and wise leadership are emulated as best as possible by the abbots of the order, who now govern in unison.\\
The mountain tops and high reaching peaks still hold much meaning for the monks of the order. The ether has not changed their mind set, it has merely allowed them to realise he is everywhere instead of just above. His dormancy and silence bothers the timeless monks not at all.

\subsubsection{Inspiration}
Tibetan monks, kung-fu movies, asian cultures, Indiana Jones, and Christianity.

\subsubsection{Detailed Unit List}

\TinySection{Apprentice}
The apprentice is a learned soul of only a hundred years or so who does his best to rise up. He is an all rounder combat unit with strike undead. His topknot symoblises he has not yet lost his pride. He bears himself humbly despite the top knot and is very sub serviant in nature. He constantly quotes the scrolls. 'Our footsteps shall drum his song'.

\TinySection{Kai Monk}
These monks have perfected the art of self defense and speed. They travel as wind fleet of foot and unhinderd. They have a suprisingly strong melee attack which is their perfect defense as their robes are not woven to turn aside blades. The are short in speech but ultimatly wise.  'I see'.

\TinySection{Healer}
This bright young lady is only five centuries old and carries herself maturley without weapon of any sort she wades into battle to administer healing and aid. She is fast of foot and can heal any wound save death. She abores violence and is a bit chirpy for an enlightend soul. 'On the battlefield all cry'.

\TinySection{Reliquery}
The holy relics of the order promote great strength of spirit and soul. Often they are carried onto the field of battle by two apprentices. The Reliquery is slow moving and without weapon or defense and therefore slows the army but gives them great strength as well. 'We carry our souls into battle'

\TinySection{Slingshot}
Some monks look to the slingshot for their attack. These monks are still strong of fist though not as strong as their kai brothers. They have poor defense and carry a massive sling on their back which needs to be deployed then shot as a direct fire weapon which deals massive damage. 'Enlightment sometimes requires a few stones to be thrown.'

\TinySection{Templar}
A vicous warroir who has been deemed unable to achieve inner peace. A large Tetsubo which stand a full three halfs her height destroys those around her with massive damage and therefore she see little need for defense, other than decoration. Her attitude is violent and slightly overconfident. 'Move aside fool'

\TinySection{Samuri}
A warroir who sees enlightment in the defense of his people and his ideals. The samuri seeks balance in all he does and seeks to achive. His strong attack is matched by his strong defense. His attitude is that of a poet. 'My blade shines in the early morning mist'.

\subsubsection{Auras}

The all embracing one is everywhere and nowhere, his voice is a whisper on the wind. As such prey will bring passive auras of strength and slow effects. Their are no magic casters in the holy city but the priests and holy men can draw strength from prey and inspire those around them.
The auras which can be generated are as follows
\begin{description}
\item [Regeneration]: Slow recuperation of health
\item [Courage]: Restores morale
\item [Honed Assault]: Improves attacks
\item [Shining Soul]: Improves defense
\item [Holy Message]: Delievers speed to the feet of warriors
\end{description}

\subsection{The Undead}

\subsubsection{Units Summary}
\begin{description}
\item [Cattle] - A slow moving sacrifical unit
\item [Virgins] - A helpless yet powerful sacrfice
\item [Maytar] - A fast devote suicide bomber, sacrifice

\item [Hopping Eyeball] - A fast scout with the jump ability

\item [Skeleton Trooper] - A cheap melee combat unit
\item [Skeleton Archer] - A accurate short ranged unit

\item [Zombie] - A slow yet powerfull melee unit
\item [Anathema] - A mass of corpses which create a powerful melee unit

\item [Necromancer] - A female caster who summons skeletons
\item [Vodoo] - A female caster who summons forth zombies
\item [Lich] - A powerful skeletal mage who has mastered the undead arts
\end{description}

\subsubsection{History}

The body is not the property of an individual it is the whole. The undead society of the ever sinking swamps believes in the never changing truths of the world. An elite few rule over the mass who live comfatable and pleasent lives for the most part. The first life is thier own. After that at death it is weighed up how much the person still owes society by the courts. The person then serves their indebted years as a automaton until his or her dues are paid in full at which point the soul is released unto the final rest. The Lich King ruled all with a fair and just hand, the society bloomed as all paid theirs dues in the end.\\
Then came the breaking. The society reeled as the courts fell and the Lich King vanished from sight. The courts eventualy were restablished but the courts are ashamed to admit a few rouge necromances and reanimators roam the lands. The few the courts took which was very popular was that corpses reanimated in battle were prisoners of war and that the would be released upon agreement, treaty, or peace. Until such time they were treated as an indebted soul.\\
The undead cannot truly understand the revulsion some races hold for their scientific art. They see an automaton as a second life and nothing more.

\subsubsection{Inspiration}

Many undead novels, dracula and classical undead sources. The culture is similiar to a the Maat of certain egyptian novels. 

\subsubsection{Detailed Unit List}

\TinySection{Cattle}
The slow plodding dumb spawn of the breeding tanks are seen as cattle despite their slightly humanoid appearence they are viewed as animals. They are given no weapons and can be mass produced cheaply. They are made to enhance their death and thus they return double life points on their death. The voice and attitude is slow and inhuman in its nature. 'Fooood'

\TinySection{Virgins}
Equiped with no weapons and only is protected by her virtue this white clad virgin has lived a pure and unspoiled life. The Virgins of white never serve a second life as their entire debt has been more than paid during this life and they are viewed as magical pure beings. Upon their death releases a massive amount of pure life force. Their attitude is selfless beauty which has a shimmering and silver quality to it. 'I serve the lich'

\TinySection{Martyr}
An extremly fast suicide bomber who has dedicated his death to the greater cause as opposed to his life. He has little armor and his defense is low but he does release a moderate amount of life force as he explodes and destroys the undisciplined fools outside the circle of life. His attitude is slightly crazed and demeted, he mentions death in almost every phrase. 'I am one with death'

\TinySection{Hopping Eyeball}
A severed eyeball which snakes, slithers and hops across the terrain. Once a turn this weaponless unit can make a massive leap, crossing obstacles or reaching great heights. The attitude of the eye is simply describe as squelch. 'Squelch'

\TinySection{Skeleton Trooper}
A simple old corpse which moves at a decent pace and has a medium melee attack. The armor is of a poor quality but the unit is the cheapest combat unit of the undead. The attitude is dry and cynical humour. 'I never know where to stop brushing my teeth'

\TinySection{Skeleton Archer}
A old skeleton archer with accurate short range who moves at a decent pace and weak defense. The skeleton is very much a whinner. He constantly complains about the food, and lack their of. 'The army chef wouldn't give me soup, I like soup'

\TinySection{Zombie}
A slow plodding unit with resilent health and a powerful attack. The unagile unit has a decompsing look about it. The attitude is forgetful and constantly inquiring if it has the correct amount of limbs. 'I knows not where my nose are'

\TinySection{Anathema}
A powerful tank of a unit created by sewing several corpses together. The dimwitted and slightly slow unit has extreme damage and is very tough. The anathema is always in several minds and argues constantly with itself. 'We need a diet! No its puppy fat'

\TinySection{Necromancer}
A gaunt lady of the night who wields a weak kris knife in one hand and a tomb in the other. She summons skeletons to serve the Lich lords and can regenerate the undead troops. She is however frail and not taken to combat, but she needs to be amoung the fray to be the most effective. Her attitude is superoir and eltist, as she glides over the terrain. Her voice cold and uncaring. 'You there, fetch.'

\TinySection{Voodoo}
Holding her voodoo totem, which allows a weak melee attack. She is vunerable as she wades into battle unable to duck due to her standing, and elite status. She reanimates the corpses of the land to the fleshly zombies and Anathemas of the undead swarm. A rich southern accent complements her dark skin and bright clothing. 'Dis her be my job'.

\TinySection{Lich Lord}
The masters of the undead who can bring forth any undead art, and judge anyone's service. These terryfying skeletal lords have permentaly bound themselves to the service of the greater and have forever cast aside the final rest. As a result they are doubly close to death and can absorb life force much easier. Their attitude is as cold as a crypt as they are judge, jury, executioner and slave master. 'I judge you unworthy to die'.

\subsubsection{Fear}

All undead units have a fearsome value this is how terryfing the other races percieve them to be. The mere presence of undead causes morale damage. The resulting effects can be panic, confusion, paralysis, and outright retreat.

\subsubsection{Death Magic}

For every unit that falls more shall rise. The undead magic users draw strength from the power of the death. The amount of power is determined by the units AP and proximity to the mage. However because of the fraility of undead mages being in the thick of battle is dangerous for them. Also any unit killed by death magic does not release mana.
Spells are in three areas.

\TinySection{Reanimate}
\begin{itemize}
\item Hopping Eyeball
\item Skelton Trooper
\item Skelton Archer
\item Zombie
\item Anathema
\end{itemize}

\TinySection{Vodoo Curse}
\begin{description}
\item[Ghoul Touch]: Damage target and steal life force
\item[Cripple]: Target looses half time units
\item[Despair]: Destroys morale
\item[Taint]: Makes an area of terrin or body of water damaging to the living
\end{description}

\TinySection{Necromancer Boosts}
\begin{description}
\item[Haste]: fifty percent increase in time units
\item[Rage]: Target unit deals double melee damage
\item[Regen]: Fully heals undead unit
\end{description}

\subsection{The Mages}

\subsubsection{Units Summary}

\begin{description}
\item[Scholar] -  One spell melee unit
\item[Barrister] - Weak vulgar magic with poor melee (dagger)
\item[Conjurer] - Weak summon magic with poor range (dart gun)
\item[Magister] - Strong Vulgar magic with no weapon
\item[Summoner] - Strong Summon magic with no weapon
\item[Eldar] - Magic Master with no melee, very weak
\end{description}

\subsubsection{History}
The desert brings forth much time for thinking and science and as such the mages of the desert long ago harnessed the magical energies of their world. They built great cities which became oasises in the desert.\\
The arch mage ruled all as the experiments and great workings of magic rand throughout the country like sand in the hour glass. The death of the arch mage during the breaking was a terrible loss for all magical arts.\\

\subsubsection{Inspiration}
Arabian, magic, fantasy, genies, alladin, djinns and Egypt.

\subsubsection{Detailed Unit List}

\TinySection{Scholar}
A student of the arcane arts who has just started to dable in his first spells and hasn't het lost his physical strength to dark deals and trade offs. He carries a large oaken staff whose shoden ends pack a large punch. Sadly he wields only a single spell. His ambition and keeness show in his young voice. 'Show me the way master.'

\TinySection{Barrister}
The first order of vulgar magic, a gauntish figure who has began to learn the arts of the elements. His border has weakend allowing him only to carry a small dagger. His voice is soft and sure as he walks the paths of power. 'Need a light.'

\TinySection{Conjurer}
The first summonings have been cast and he knows the binding symbols. His body weakens as he carries a small dart gun. He begins to summon creatures of power, one at a time. His voice is loud and sure as he commands his minoins. 'Bow before your new master'.

\TinySection{Magister}
His body has faded past all physical tasks but he has master the vulgar magics and has the very elements flowing through his veins. He throws fireballs and lightning arcs between his footsteps. His voice is a whisper of silent malice. 'I command the elements'.

\TinySection{Summoner}
The ancient tombs of bindings and incantations of control are known to him as his withered form commands contructs and elementals of great power. He can command two creatures at any given time. His voice is clear and unwavering, echoing in all minds. 'Heel'.

\TinySection{Eldar}
The physical has faded until only a thin shell of life remains, blood replaced with pure power. The magics all mastered and the channels open for combat. None stand in his way. His voice is like ten voices at once. 'I am magic.'

\subsubsection{Art of Magic}
The mage team has several casting units but they all use this system. The mage fills a spell book with spells. Each spell has a power rating. The mage is limited to selecting either a broad range or a specialised range. For example a mage could take 8 power rating 1 spells or one power rating eight spell, or a mixture.
The mage starts with a set amount of mana, there are units that can help recharge mana for a mage. The mage can cast any spell numerous times but each time he expends mana. The mage may cast as many spells as his TUs and mana allow.
The spell list is too diverse to list at the moment but the magic falls into three catergories:

\TinySection{Vulgar}
These are direct Effect spells which attack a certain target with a certain effect.
\begin{description}
\item[Ice Bolt ]: Low Damage, reduce TUs next turn
\item[Fire Bolt ]: Weak Damage (better than Low Damage)
\item[Lightning Bolt ]: Low Damage travels on

\item[Freeze Ball]: Explosion which freezes targets
\item[Blaze Ball]: High damage explosion
\item[Sonic Boom]: Damaging explosion which stuns

\item[Ice Shard]: Damages and freezes target
\item[Lava Blast]: Exteme direct damage
\item[Chain Lightning]: Moderate damage that jumps until it can no longer

\item[Haste]: Increase TUs for target, next 2 turns.
\item[Time Speed]: Increase Tus for team, next turn.
\item[Slow]: Decrease TUs for target, next 2 turns.
\item[Bullet Time]: Decrease TUs for team, next turn.

\item[Teleport]: Jump to any visible point on the map
\item[Recall]: Recall any ally to your side.

\item[Sonar]: Detect the location of all foes.
\item[Feather Flight]: Fly for a turn.
\end{description}

\TinySection{Summon}
These spells bring forth entities to fight for the mages. Other spells bring forth shields and barriers.

\begin{description}
\item[Minor Elemental]
\item[Elemental]
\item[Magor Elemental]
\item[Sand Golem]
\item[Iorn Golem]
\item[Rock Golem]
\item[Wall of Fire] : Creates a damaging wall of fire
\item[Air Wall] : Creates an imprentrable wall for one turn.
\item[Ring of Fire] : Creates a damaging ring of fire
\item[Dome of Air] : Creates a invunrable dome for one turn
\item[Elemental Armor] : Protects target
\item[Bridge of Air] : Creates a bridge of air
\item[Minor Portal] : Creates a protal for one turn to visible location.
\end{description}

\TinySection{Strategic or Universal}
These are indirect spells which give the mage a long term stratgic adcantage. All mages can use these spells.
 
\begin{description}
\item[Mage Armor] : Increases defense for the mage only
\item[Far Sight] : Looks at a small area of the map
\item[Blink] : Random short range teleport
\item[Mana Strike] : Damaging melee attack
\item[Mana Arrow] : Damaging ranged attack
\item[Leap] : Allows a long or high jump.
\end{description}

\subsection{The Alchemists}

\subsubsection{Units Summary}

\begin{description}
\item[Spug] - Basic very fast weak short range unit with one potion.
\item[Spider] - Very fast all terrain unit with two potions.
\item[Burrower] - Slow underground unit with two potions.
\item[Potion Master] - Fast weak melee unit with lots of potions.
\item[Sniper] - Long range good tunnel vision with one potion.
\item[Mortar Man] - Long range indirect fire, can launch potions, a few potions.
\item[Glider] - Flying unit with a few potions and no melee.
\end{description}

\subsubsection{History}

The cavern world of the Alchemists was like a swiss chessball riddled with caves and tunnels and occansional exploding. They lived in a world where every goblin was his own and piracy, corruption, crime, pranks, and plain stupidity are the orders of the day. The goblins are personified by an innate lack of caution, planning or discreation.\\
They brew magical potions in complex laboratories and in truth the closest they've ever had to a goverment or ruling body is the Council of Spoink every year, which is mostly an excuse to get drunk. They claim to need no goverment as each goblin is intelligent enough to rule himself, the truth is more likely to be along the lines of no goblins bloody stupid enough to try tell other goblins what to do.\\
When the breaking occured the goblins were most probably the least phased, the biggest shock was discovering they didn't have a roof over their heads, the subterrain race quickly adapted however as they always do. The only word close to tradition in their native tounge is Juu'roek which means old broken idea. They were also the first race to discover portals, and many other things mostly through blind dumb luck. This philosphey has severed them well in the past and they seem to be sticking by it.

\subsubsection{Inspiration}

Goblins, pirates, fantasy novels, alchemists.

\subsubsection{Detailed Unit List}

\TinySection{Spug}
Basic warrior, scout, lab rat and general tool. The spug carries a single potion and a stack of javelins which are extremely short range and rather weak. His armour is none exsistent. HE is extremly fast on foot however and has twice the normal climb speed. He carries only a single potion. His attitude is inquisivtive and rather squeeky. 'Wanta see me go squiiiish!'

\TinySection{Spider}
The all terrain spider goblin climbs vertical terrain at extreme speeds almost negating their effect. He is also capable of a short jump. He has the same small stack of weak javelins as a Spug but carries two potions. His armour is weak but his swift navigation of the terrain gives him the advantage over the slower and more gravity bound teams. His attitude is that of a hyper active six year old on a sugar rush on christmas eve. 'Spidy wanna jump, jumpie jumpie!'.

\TinySection{Burrower}
This dirty little rat goblin dives under the ground and digs at a relativly fast pace. Inflitrating behind enemies and dropping one of the few potions he carries. He also has a weak melee attack. Going from above ground to below or visa vesa however takes most of his time units. His attitude is that of a french underground resistance fighter, on helium. 'Vey vill not see me'.

\TinySection{Potion Master}
This weak little old fool with a puny weapon and even worse armor has the most devasting large potion belt of all. Allowing him to carry a large amount of potions into battle. He is still fast like all other goblins however. His attitude is that of a demeted genius who snapped around the bend too many years ago to count. He could be compared to mad professor or university lecturer. 'Oh it might blow up before you throw it but thats what the buckets are for.'

\TinySection{Sniper}
This posh know it all goblin has a rather powerful range attack that takes him a full turn almost to deploy, but once deployed the fast defenless climber, with no weapon becomes a far sighted eagle with tunnel vision and brilliant range. He carries a potion on him at all times and can throw them while deployed. His attitude is a soft spoken hot potato accent which is sure of his aim. 'I see them they die.'

\TinySection{Mortar Man}
Long indirect range which is rather devastating, even if extremly inaccurate, area of effect damage. They can also launch their potions via the motar cannon, of which they carry a few. They are fast as all goblins and almost unarmoured, they climb at double speed. They have a rather disturbing habit of referring to everyone as sarge and sound as if they have been stoned for the entirity of their lives. 'Sarge was we meant to hit dat.'

\TinySection{Glider}
This terror of the skies is a fast extremly fast climber with a hang glider on his back, when sober at all other times he is a menance to society. Once this ill armoured unit with no weapons reaches high ground he jumps off and flies for a while all the time dropping potions and small rocks on unsuspecting victims. When he is desperate he can launch into action but it takes a full two turns to launch from a low altitude point. He is convinced of his piratedom and constantly uses foul langauge. 'Oi Lass guess what I can see'.

\subsubsection{Expendable}

The brewers of potions, magic dusts, and exploding toys. Alchemy is an art which cannot be praticed on the field of battle. However, the potions and dusts concoted can be carried into battle. All potions have an AP value.\\
Almost every goblin on the field of battle has a potion belt which holds the potions of his people. He uses them much like grendades. Whenever a goblin dies there is a chance of his potions going kabooom! Some potions are listed below
\begin{description}
\item[Popper]: Small explosion
\item[Bang Bottle]: Explosion
\item[Kaboom]: Biiig Explosion
\item[Gloop]: AOE which sticks, prevents movement, for all units for a turn in AOE.
\item[Slip Slap Sponk]: An invisble liquid which makes a small slick which will cause the next walking unit to cross it to fall and loose all his Time Units
\item[Feel Ills]: An AOE poison
\item[Party Spice]: AOE which causes confusion for a turn.
\item[Scream Juice]: AOE Detroys Morale
\item[Rock Goo]: AOE creates a big hole
\item[Sparkle Dust]: AOE Partial Blinds units for a turn.
\end{description}
 
\subsection{Nature}

\subsubsection{Units Summary}
\begin{description}
\item[Fairy] - A female scout unit with the ability to hover
\item[Druid] - A male support caster unit

\item[Elven Archer] - A male elf with bow

\item[Elven Ranger] - A male elf with dagger
\item[Dryad] - A female range unit 

\item[Nymph] - A female spell caster
\item[Ent] - Strong melee unit
\end{description}

\subsubsection{History}

Through melody of the God of Nature Ethos , Etharia was created. The lands were lush and green, covered with beautiful forests and rich plains. When the creatures of Etharia came to life, they rejoiced in the beauty of Mother Nature, cherishing the gifts of beauty that came from her. These creatures of nature, possessing very patient temperaments, took from nature only what was needed, leaving plenty for others to use. When the jewel cracked and chaos reigned, the darkness crept into the land of nature through the mysterious portals, and the harmony of Etharia was disrupted, Nature's gifts and resources began to be abused. The nature folk still cherished their forest, choosing to build homes from stone, or fallen logs. The darkness that now reigned in the hearts of the wicked turned them green with jealousy, and thus they attacked the forests. Terrible fires were set, and entire forests disappeared, leaving nothing but barren sands in its stead. Although the Nature folk do not like to get involved in politics, their pain of losing their home goaded them into tracking the malevolent creatures and disposing of them, thus bringing peace and protection to their beloved lands. 

\subsubsection{Inspiration}
Fairytales, fantasy novels, Welsh Myth \& Tolkien Fantasy

\subsubsection{Detailed Unit List}

\TinySection{Fairy}
These miniscule creatures, benevolent in nature, are care-free and easygoing . They enjoy roaming the lands in search of a good time. Although a scout unit with no weapon, they do have enthralling spells and good eyesight radius. 'Fairy dust tickles! hee hee'

\TinySection{Druid}
The tribal guardian that will stop at nothing to end the encroachment of civilisation into his lands, the Druid stands apart from all others in the fantasy setting as the most mysterious and versatile of individuals, dedicated to principles lesser creatures do not understand and reaping power others can only gasp at. The mystery and majesty of the druid, his rich and versatile career all make for a great support caster be it a weak melee.'Let's go skinny dipping!'

\TinySection{Elven Archer}
Elves are an alien and rare race, when compared to the other races.  They are fast in their movements and thoughts. A more obvious attribute being their pointy ears, which is how many races refer to them. Elves are very graceful in stature and extremely proud, they tend to believe that they are far superior than any other creature. Which is not a far-fetched assumption since they are highly-intelligent, they can learn skills faster than any other race. Being able to concentrate on a single task without letting up. And far surpass other races in song, writing, magic, architecture, or any other fine art. As for immortality, although they can live for a very long time as they do not really die naturally, they can be slain in battle.Elven Archers tend to be tall, thin and usually no stronger than a man. They are good range units with good movement. No Melee.'No my name isn't Legolas!'

\TinySection{Elven Ranger}
Known as the Lifeblood of the Forests, the Rangers are the guardians and keepers of Nature. Hidden within the forests, they guard the trees and lands of Elves, welcoming all friends and disposing swiftly of all enemies. The Ranger has the ability to survive in a hostile environment on their own with little supply, thus they are invaluable in rescue parties and hunts. Their ability to track and detect movement of creatures gives them the advantage of being able to catch any threat that dares walk through the Elven lands. The special bond that the Ranger has with Nature not only helps them protect her, but provides them with Nature's protection as well.An excellent range unit with good eyesight radius, poor defense and a dagger for a weapon. 'He who speaks little, wise is he'

\TinySection{Dryad}
The Dryads are scantily clad, exquisitely beautiful female spirits of the oak trees that are so charming that they can entice their victim to do whatever their heart desires. The Dryads are good range units with magic abilities. 'Do you come by this side of the forest often?'

\TinySection{Nymph}
The nymphs are also scantily clad, exquisitely beautiful female spirits of nature, who though living many years are nevertheless bound to die; some talk however as if they were immortal. The nymphs can be aggresive in their love and also extremely seductive. They have the divine power of changing the shape of things: some women were transformed into nymphs by the nymphs themselves. These spell casters have a medium range. 'Hey there handsome'

\TinySection{Ent}
As the tree herder of the forest, the Ent has to protect his beloved trees.  The Ent is a huge, powerful melee unit that fires massive boulders at attackers indirectly from great distance,  thus causing great damage wherever it falls. The Ent is a loyal and diligent worker that is very passionate about its calling. 'I need to get a sheepdog er... make that a treedog?!'

\subsubsection{Life Magic}

\TinySection{Recharge}
The nature of the forest is that of patience. Every nature mage has magic which when cast will slowly refresh, on a time basis. The system allows him to use his spells in any order as long as they are refreshed.Nature mages will also have a meditate option to refresh faster while doing nothing.

\TinySection{Fairy}
\begin{description}
\item[Concealment]: Unit obtains invisibility while stationary
\item[Blind]: Target loses half time units
\end{description}

\TinySection{Druid}
\begin{description}
\item[Summon]: Grizzly Bear, Hawk, Fox
\item[Tremor]: Area effect that stuns and damages target *
\item[Bog]: Target loses half time units *
\end{description}

\TinySection{Dryads}
\begin{description}
\item[Tree Stride (Nature Stride)]: Teleport between two lands objects
\item[Entwine]: Target cannot move next turn *
\item[Bark Armour]: Increases defense *
\item[Poison]: Damages target over time * 
\end{description}

\TinySection{Nymph}
\begin{description}
\item[Water stride]: Teleports between two points in a body of water
\item[Seduction]: Target loses half time units *
\item[Heal]: Heals injured unit by fifty percent *
\item[Mist]: Area effect that reduces sight range +
\item[Encourage]: Morale increase +
\end{description}
\footnotetext{Certain spells share timers, they are indicated with a * or + note this doesn't cross over unit barriers}

\subsection{The Mechana}

\subsubsection{Units Summary}

\begin{description}
\item[Steamhead] - A gnome skilled repairing machines and stocking boilers to boost steam.

\item[Rat-a-tat-tank] - A small yet fast hovering tank with ranged attack, capable of making small jumps.
\item[Steam Fiend] - A slow but powerful tank with a fairly devastating direct blast and the ability to charge into battle.
\item[Big Boom] - A slow tracked shell launcher that provides earth-shattering artillery support, and is capable of firing a reconnaisance flare.
\item[Lumberjack] - An upright tracked machine with a spinning saw blade at the end of each of two arms, capable of clearing terrain and dealing massive melee damage.
\item[Slice-n-Dice] - A hovering cutting machine that spins horizontally.
\item[Steamcopter] - A flying machine capable of dropping bombs on the unsuspecting enemies below
\end{description}

\subsubsection{History}

Since the time that steam poured from the termal vents of the foothills as geisers, the Source has existed. There are numerous phylosophies of how the Source was compiled. Some say that deep within the crust of our home vapour, water, and stone all flowed together to create the ultimate machine. Others say the Source was the ultimate creation of another civilisation, just before they were destroyed. It is also speculated that the Source was a hybrid of smaller machines that combined pressure to rule the world. The origin of the Source is of little importance. What is important is that it dissapeared after the Breaking, and we are left to our own computations. We built the academies to train the Steamheads, the maintainers. But with the world being subject to division we have little space left to construct our plants and factories, and even less minable metal resources. And so we find ourselves in dire need of multiplication, or existance will be subtracted from us. \\
As a result in this chaotic non-sensical reality we need new machines to battle for dominance over these floating islands of rock in the endless ether. We fight moving trees, monks of stone and illogical goblins. But we remain true to the Mechana.

\subsubsection{Inspiration}

Steampunk anime, computer games, films and cartoons, as well as cyberpunk literature. It is based on the concept of advanced technologies, but all of them steam powered.

\subsubsection{Detailed Unit List}

\TinySection{Steamhead}
From the large academies of the Mechana world, the Steamheads are the elite of their class. Masters of physics, mechanics and dynamics, summarised as Steam Engineers, they are capable of singlehandedly constructing and maintaining any mechanical steam powered device, regardless of the complexity or size of any such machinery. Although not trained to parttake directly in combat themselves, they are accustomed to combat situations, and do not flinch when gears fly and boilers burst. More importantly, their presence is vital to ensure the large machinery stays operational in the heat of battle, at all times. 'Me Fixy!'

\TinySection{Rat-a-tat-tank}
The smallest and most agile of the Mechana machinery, the Rat-a-tat-tank is capable of almost stunt-like manuevarability. The hovering body is suspended on a cushion of accellerated steam projected from the base of the unit. To keep the weight down and to ensure a high speed, it is less heavily armoured than its bigger brothers, and has been fitted with a smaller barrel for the same reason. The Rat-a-tat-tank is also capable of releasing a sudden burst of steam, sending in skyward for a short distance, allowing it to overcome more cumbersome terrain. 'What be ye rattle-prattle?'

\TinySection{Steam Fiend}
What the Steam Fiend lacks in speed it makes up in armour and sheer firepower. Making use of the very best steam propulsion engine the Steamheads can muster, this unit is an engineering masterpiece. It is moved forward on huge tracks, and sports a magnificent barrel. It also has the ability to build up steam and charge into battle. 'Lets start steam-rolling!'

\TinySection{Big Boom}
This metal beast propel steampowered shells over large distances. Because of the high precision requirements the barrel has to be fairly long to maintain accuracy. Because of this the big boom is fairly slow. To avoid being swarmed however, it is equiped with a reconnosaince flare to expose the surrounding terrain.'Valves sealed and pressure feeled'

\TinySection{Lumberjack}
The lumberjack is an upright machine, and would have been a bi-ped if it were not for the tracks that propel it. Its moderate speed serves it well as it needs to get into battle faster than the slower units. It has a central cockpit, with a pilot. To each side of the cockpit an arm has been fitted. On the end of each arm is a spinning blade, lethal to anyone that would approach it. Besides ripping its enemies to pieces, the Lumberjack is also capable of clearing terrain for the heavier units.'Cut-n-Paste!'

\TinySection{Slice-n-Dice}
The Slice-n-Dice combines the state of the art steam engine of the Steam Fiend with the steam cushion the Rat-a-tat-tank. Hovering above the ground the Slice-n-Dice takes a few moments to reach optimal spin speed, but anything around it within the range of its blades will be processed meat within a fraction of a second. After its frenzy has finished it needs to cool down however, or risk blowing every single steam valve in the system. 'Optimal angular velocity reached'

\TinySection{Steamcopters}
The Steamcopter is the first mechanical unit capable of flying, and it is the pride and joy of Steam engineering. Although it is still in the prototype stage, many future designs will be based on it, hopefully giving the Mechana the upper hand. It has a spinning rotor blade at the top of its chassis. This blade only serves as a way to keep the copter aloft, and cannot be used in combat. Also, the copter can any travel through the air when it has acculumlated enough steam pressure to keep the rotor blade spinning for prolonged periods of time. It is capable of exploring the terrain over larger distances, and can even drop explosives on unsuspecting targets from a high altitude. 'Lopping me Bombies!'

\subsubsection{Abililties}
\begin{description}
\item[Repair]: This ability to repair Mechana units lies exclusively with the Steamhead.
\item[Boost Steam]: The Steamhead can stock the boiler of any unit, allowing it to build up steam pressure quicker than on its own steam.
\item[Hover]: units with hover capability float on steam at a set distance from the ground. Although this means they are not affected by terrain, they cannot make use of cover that the terrain would otherwise have provided.
\item[Jump]: The Rat-a-tat-tank can traverse small crevices, or just move forward quicker by using its jump ability.
\item[Charge]: The Steam Fiend can build up steam and charge into battle. This allows it to get quickly into a critical situation at a critical time.
\item[Reconasaince flare]: The Big Boom has the ability to launch a flare and illuminate an area on the ground, exposing units beyond the normal site range of Mechana units.
\item[Clear Terrain]: The Lumberjack is capable of clearing terrain that would have otherwise hampered the heavier Mechana units.
\item[Bombing]: The Annoyer can hurl bombs at units from a high altitude.
\item[360 attack]: The Slice-n-Dice has 360 attack. It deals damage to all units in adjacent squares.
\end{description}

\end{document}
